Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being

### 2023-04-12 outputs/paper/paper.qmd
Minor changes

---
title: "Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being"
subtitle: "A Cross-Sectional Cumulative Analysis of U.S. Microdata from 1972-2012"
author: 
  - SHAOHAN CHANG
thanks: "Code and data are available at: https://github.com/lucas11333/Of_Wellbeing"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This study analyzed cross-sectional cumulative microdata from the United States spanning 1972 to 2012 to investigate the impact of family size on happiness, health, gender dynamics, and financial well-being. The number of children served as a proxy for family size, allowing the examination of relationships between happiness, health, and financial satisfaction in families. Findings indicate that childless individuals experience the highest happiness levels, and one-child families exhibit the best health. Furthermore, increased financial satisfaction correlates with fewer children within a family. These insights hold significance as they enhance our understanding of the factors influencing family size and the potential consequences of family planning decisions on overall well-being."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This following data analysis is processed in R [@R] with packages of tidyverse [@ti], dplyr [@dplyr], janitor [@janitor] , knitr [@knitr], here [@here], haven [@haven], ggplot [@ggplot] , kableExtra [@kableExtra] , model summary [@modelsummary] to assist in data analysis, and the data from General Social Survey [@gss].

@@

The following table shows the first five rows of data and their corresponding parameters.The following data section provides a more detailed explanation of the selected variables for the data.

```{r}
#| echo: false
#| warning: false
library(tidyverse)
library(dplyr)
library(tidyr)
library(kableExtra)
library(janitor)
library(knitr)
library(haven)
library(here)
library(ggplot2)
library(tinytex)
library(modelsummary)
data <- read_csv(here::here("inputs/data/cleaned_data.csv"))
```

In @tbl-var contains information on various personal and demographic characteristics of individuals, collected between the years 1972 and 2012. The data-set comprises of 10 variables, namely age, sex, number of babies, divorce status, self-reported health status, financial satisfaction, level of happiness, and additional variables. The data-set contains 38,960 rows, providing ample opportunities to analyze the relationships between these variables and gain valuable insights into the factors that could potentially impact an individual's overall life satisfaction.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-var
#| tbl-cap: Report Data (First 5 rows)
data %>%
  select(1:6) %>%
  slice(1:5) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```{r}
#| echo: false
#| warning: false
data %>%
  select(7:10) %>%
  slice(1:5) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

-   year: This variable indicates the year in which the data was collected from GSS. In this data-set, all the observations are from the year of 1972 to 2012.
-   sex: Categorical variable with two levels: male and female.
-   babies: Integer variable with different levels, depending on the number of babies, for instance : 0, 1, 2, 3, 4, or 5+.
-   health: This variable indicates the self-reported health status of the individual at the time. It is a categorical variable with different levels, depending on the health status, for example : excellent, good, fair, or poor.
-   finance_satisfied: This variable indicates the individual's different level of satisfaction with their current financial situation. It is a categorical variable with different satisfaction levels: very satisfied, more or less satisfied, not very satisfied, and not at all satisfied.
-   happy: This variable indicates the individual's level of happiness or life satisfaction. It is a categorical variable with different levels of happiness: very happy, pretty happy, not too happy.
-   happy_score: The happy_score variable is an integer that represents the individuals in the dataset's self-reported happiness levels. Its sign ranges from 0 to 3, with 0 denoting "not too happy," 1 denoting "happy," 2 denoting "very happy," and 3 holding any unclassified or missing responses.
-   health_score: The health_score variable is an ordinal variable shows the self-reported health status of the individuals, which sign the ranges from 1 to 4. With 1 denoting "poor" health, 2 signifying "fair" health, 3 representing "good" health, and 4 referring to "excellent" health. The value 5 is reserved for any unclassified or missing responses.
-   finance_score: This ordinal variable represents the financial satisfaction level of the individuals in the dataset. It ranges from 1 to 3, with 1 indicating "Not at all satisfied," 2 denoting "more or less satisfied," and 3 signifying "pretty well satisfied." The value 4 is reserved for any unclassified or missing responses.
-   babies_number: The babies_number variable is an ordinal variable denoting the number of babies an individual in the dataset has. It ranges from 0 to 4, where 0 represents "0" babies, 1 denotes "1" baby, 2 signifies "2" babies, 3 indicates "3" babies, and 4 refers to "4" babies. The value 5 is reserved for any unclassified or missing responses.

@@

In [@fig-one],which analysis explores the relationship between the number of children in a family and happiness levels, categorized as 'very happy,' 'pretty happy,' and 'not too happy.' Although childless families show the highest numbers in each happiness category, the data collected for such families is more abundant, necessitating an analysis of happiness proportions within each family size. The findings reveal no linear correlation between the number of children and overall family happiness, with happiness varying across family sizes without a consistent pattern. The non-uniform distribution of happiness categories among families with different numbers of children implies that factors beyond the number of children also contribute to happiness levels. In conclusion, the data shows variations in happiness levels among families of different sizes but does not establish a definitive correlation between the number of children and overall family happiness. Further research is needed to better understand the complex factors influencing happiness within families.

```{r}
#| label: fig-one
#| fig-cap: Relationship between Number of Babies and Happiness
#| echo: false
#| warning: false
data %>%
  group_by(babies, happy) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(babies), y = prop, fill = happy)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(name = "Number of Babies", expand = c(0.1, 0.1)) +
  scale_y_continuous(labels = scales::percent_format(), name = "Percentage", expand = c(0.05, 0)) +
  scale_fill_discrete(name = "Happiness") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
        plot.subtitle = element_text(size = 14, margin = margin(b = 20)),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")

In [@fig-two], as presented in the bar chart, the gender distribution varies depending on the number of babies in the family. By comparison, a higher percentage of females than males is consistently observed across all groups. On the other hand, the lowest difference in sex distribution is found among families with no babies, where the female-to-male ratio is relatively closer. As the number of babies in a family increases, the percentage of females increases, with the highest difference observed in families with one to four babies. In these groups, the sex distribution is more skewed towards females.

```{r}
#| label: fig-two
#| fig-cap: Relationship between Number of Babies and Gender
#| echo: false
#| warning: false
data %>%
  group_by(babies, sex) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(babies), y = prop, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(name = "Number of Babies", expand = c(0.1, 0.1)) +
  scale_y_continuous(labels = scales::percent_format(), name = "Percentage", expand = c(0.05, 0)) +
  scale_fill_discrete(name = "Sex") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
        plot.subtitle = element_text(size = 14, margin = margin(b = 20)),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")

In [@fig-three],which shows analysis and visualization of the number of children born each year between 1972 and 2012 in connection to the degree of happiness to offer a more in-depth and complete perspective of the data. In 1974, 1985, 1993, 2004, 2006, and 2008, a relatively high percentage of people reported feeling unpleasant concerning the variable "not too happy." On the other hand, the proportion of people reporting being cheerful in 1973, 1977, 1994, 2006, and 2012 was relatively high. The data shows that excellent, fair, and sound are the most prevalent self-rated health statuses among adults with one or two children. It cannot be claimed that the number of children will impact the health scores, even though participation rates in families with one kid are generally high, even as time goes on.

@@

ggplot(data[data$happy == "very happy", ], aes(x = year, fill = factor(babies))) +
  geom_bar(position = "fill") +
  facet_wrap(~health, ncol = 5, scales = "free_y") +
  labs(x = "Year", y = "Percentage of Babies", fill = "Number of Babies") +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0.05, 0)) +
  theme(panel.grid.major.x = element_line(color = "grey", size = 0.1),
        strip.text = element_text(size = 10),
        panel.spacing = unit(0.2, "lines"))

@@

In [@fig-four].When individuals have one child, more participants rate their health as 'good' and 'excellent' compared to other self-health rating categories. By contrast, for the individuals who do not have children, the proportion of those who rate their health as "poor" is the highest; however, no clear conclusion can be drawn from this finding. For individuals with three children, the feedback on their health status is evenly distributed among "good," "fair," and "excellent," with a tiny proportion of participants reporting "poor" health. Overall, having one child is associated with most participants reporting an excellent health status.

@@

In [@fig-five]. The association between health and the number of babies in the family is investigated via a dot plot.  The information is divided into five categories depending on the number of babies, which from 0 to 4, and four categories based on health, including "excellent," "fair," "good," and "poor." from the year of 1970 to the 2010. The graph shows a trend that may point to a link between a family's health condition and the number of children they have. By comparison; with more kids are born to families with "excellent" health than to those with 'poor' health, with resulting in a more balanced distribution of births. As the result; this pattern suggests that having one or more children is more likely when one is in better health self-rate. The percentages for all four health categories tend to decline as the number of newborns rises. This trend is most apparent in the "poor" health category, where the percentage drop is more dramatic as the number of births rises. The information points to a connection between family size and health. The chance of having one or more children is often more significant when one is in better health than when one is in worse health. But it's important to remember that the link could not be direct and that other variables might also affect how many children a family has.

@@

ggplot(data[data$happy == "very happy", ], aes(x = year, y = factor(babies))) +
  geom_point(position = position_jitter(width = 0.2, height = 0.1), size = 2, color = "#4daf4a") +
  facet_wrap(~health, ncol = 5, scales = "free_y") +
  labs(x = "Year", y = "Number of Babies") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.grid.major.x = element_line(color = "grey", size = 0.1),
        strip.text = element_text(size = 10),
        panel.spacing = unit(0.2, "lines"))

@@

In [@fig-six]. The graph suggests the connection between the number of babies and self financial satisfaction rate. Regardless of the group, the less financially satisfied families having one or more children.On the other hand; most families in all financial satisfaction levels have 0 children, which is intriguing. In contrast to the more or less content and 'not satisfied at all' categories, the proportion of families with no children is much higher in the 'very well satisfied' group. As the number of babies grows, the percentages for all three types of financial satisfaction tend to decrease. This trend is evident in the "pretty well satisfied" category, where the percentage decrease is steeper as the number of babies increases. In conclusion, the results point to a relationship between family size and financial contentment. Families that are more financially satisfied tend to have fewer children, whereas those who are less financially satisfied tend to have more. Remember that this link could not be linear and that other variables might affect how many children a family has.

```{r}
#| label: fig-six
#| fig-cap: Relationship between Happiness and finance satified
#| echo: false
#| warning: false
ggplot(data, aes(x = finance_satified, fill = factor(babies))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Rate of Finance Satisfaction",
       y = "Percentage of Cases by Number of Babies",
       fill = "Number of Babies") +
  theme_minimal() +
  coord_flip()

# Model : Linear Regression

@@

In conclusion, the data analysis presented in [@fig-one] to [@fig-six] reveals several key findings related to the number of children in a family and its association with happiness, health, and financial satisfaction. Overall, the families without children reported that the highest happiness levels, however; the larger sample size of childless families could influence this result. The data on gender distribution shows that there are more women in families with more children compared with male. The most significant differences between families with one child and those with four children can be seen.

@@

## Table of Relationship between Number of Children and Gender.

In the @tbl-sex, which presented in the table indicate that the sex distribution varies depending on the number of babies in the family.Across all groups, there is a consistent pattern of a higher percentage of females compared to males. The lowest difference in sex distribution is observed among families with no babies, where the female-to-male ratio is relatively closer 54.53% female vs. 45.47% male. As the number of babies in a family increases, the percentage of females also increases, with the highest difference observed in families with 1 and 4 babies. In these groups, the sex distribution is more skewed towards females, with 60.92% and 60.98% female representation, respectively.

```{r}
#| label: tbl-sex
#| tbl-cap: Relationship between Number of Children and Gender
#| echo: false
#| warning: false
sex_table <- data %>% 
  group_by(babies, sex) %>% 
  summarize(response = n(), variance = var(as.numeric(sex))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, sex, percent) %>%
  spread(sex, percent)

kable(sex_table  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

## Distribution of Happiness Levels Across Families with Different Numbers of Babies

In the @tbl-happ, which suggests that families with 4 babies have the highest proportion of "happy" individuals at 63.41%. In contrast, families with 3 babies have the highest proportion of "not too happy" individuals at 16.57% and the lowest proportion of "happy" individuals at 50.30%. Families with 0, 1, and 2 babies have relatively similar distributions of happiness levels. It is worth noting that families with 3 babies have the highest proportion of "very happy" individuals at 33.14%, followed closely by families with 2 babies at 32.95%. Families with 4 babies have the lowest proportion of "very happy" individuals at 24.39%.

```{r}
#| label: tbl-happ
#| tbl-cap: Happiness Levels Across Families with number of Babies
#| message: false
#| echo: false
#| warning: false    
happy_table <- data %>% 
  group_by(babies, happy) %>% 
  summarize(response = n(), variance = var(as.numeric(happy))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, happy, percent) %>%
  spread(happy, percent)

kable(happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

## Distribution of Self-Rated Health Status Across Families with Different Numbers of Babies

In @tbl-heal, which appears suggests that individuals without any babies tend to have higher percentages in all health categories, especially in the "poor" health category (93.52%). As the number of babies in a family increases, the proportion of individuals in the "excellent," "fair," and "good" health categories generally decreases. The "poor" health category also follows this trend, but the decrease is less pronounced. Individuals with 1 or 2 babies have relatively similar health status distributions, with slightly higher percentages in the "excellent" and "good" health categories compared to the "fair" and "poor" health categories. Individuals with 3 or 4 babies have the lowest percentages across all health categories.

```{r}
#| label: tbl-heal
#| tbl-cap: Health Status with number of Babies
#| message: false
#| echo: false
#| warning: false    
health_happy_table <- data %>%
  group_by(health, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(health, babies, percent) %>%
  spread(babies, percent)

kable(health_happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In @tbl-fin, which appears suggests that families with 0 babies have the highest percentages of financial satisfaction across all categories. As the number of babies in a family increases, the proportion of families in each financial satisfaction category generally decreases. Families with 4 babies have the lowest percentages in all financial satisfaction categories.Notably, the "pretty well satisfied" category shows the steepest decline in percentages as the number of babies increases.Families with 0 babies in this category have the highest percentage of financial satisfaction at 85.47%, while families with 4 babies have the lowest percentage at 0.05%.

@@

The analysis of the GSS data showed in [@fig-one] and [@fig-five], which suggested the relationship between the number of children in a family and the level of the happiness and health status of individuals. In compare; the families with no children seem to have the highest proportion of 'very happy', 'pretty happy' responses rate. Although this could be influenced by the larger amount of data collected for childless individuals. On the other hand, the families with 'excellent' health tend to have a more balanced distribution of children, with a higher likelihood of having one or more children compared to families with poorer health.

In addition, the relationship between family size and happiness, as well as family size and health. The data analysis in [@fig-two] also indicated the relationship between family size and gender distribution. This paper suggested that the sex distribution varies depending on the number of babies in the family, with a higher percentage of females than males observed across all groups. As the number of babies in a family increases, at the same time, the percentage of females also increases, with the highest difference observed in families with one and four babies. By comparison, in these groups, the gender distribution is more skewed towards females. This finding suggested that the presence and number of babies in a family may also influence the gender composition of the family.

On the other hand, the data from [@fig-three] and [@fig-four] supports the relationship between family size, happiness, and health. The analysis of the number of children each year from 1972 to 2012 about the level of happiness shows the proportion of individuals who reported feeling "not too happy" and "happy" over the years. This finding highlights the potential influence of horticultural and economic factors on happiness and family size over time.

Additionally, the data from [@fig-four] shows that families with one child are associated with the highest proportion of participants reporting 'excellent' health status. This finding might suggest that having a smaller family size could be related to better health outcomes for individuals. However, it is crucial to consider other factors that could influence the health status of individuals, such as access to healthcare, lifestyle choices, and overall well-being.

@@

## Weakness:

@@

## Next Steps:

@@

NAME : SHAOHAN CHANG Institution/Department : Department of statistics (University of Toronto) Email Address : shaohan.chang\@utoronto.ca

>>>>

---
title: "Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being"
subtitle: "A Cross-Sectional Cumulative Analysis of U.S. Microdata from 1972-2012"
author: 
  - Shaohan Chang
thanks: "Code and data are available at: https://github.com/lucas11333/Of_Wellbeing"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This study analyzed cross-sectional cumulative microdata from the United States spanning 1972 to 2012 to investigate the impact of family size on happiness, health, gender dynamics, and financial well-being. The number of children served as a proxy for family size, allowing the examination of relationships between happiness, health, and financial satisfaction in families. Findings indicate that childless individuals experience the highest happiness levels, and one-child families exhibit the best health. Furthermore, increased financial satisfaction correlates with fewer children within a family. These insights hold significance as they enhance our understanding of the factors influencing family size and the potential consequences of family planning decisions on overall well-being."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This following data analysis is processed in `R` [@R] with packages of `tidyverse` [@ti], `dplyr` [@dplyr], `janitor` [@janitor], `knitr` [@knitr], `here` [@here], `haven` [@haven], `ggplot` [@ggplot], `kableExtra` [@kableExtra], `modelsummary` [@modelsummary] to assist in data analysis, and the data from General Social Survey [@gss].

@@

This section provides a more detailed explanation of the selected variables for the data.

```{r}
#| echo: false
#| warning: false
library(tidyverse)
library(dplyr)
library(tidyr)
library(kableExtra)
library(janitor)
library(knitr)
library(haven)
library(here)
library(ggplot2)
library(tinytex)
library(modelsummary)

data <- read_csv(here::here("inputs/data/cleaned_data.csv"))
```

@tbl-var illustrates the nature of the data-set. The data-set contains information on various personal and demographic characteristics of individuals, collected between the years 1972 and 2012. The data-set comprises of 10 variables, namely age, sex, number of babies, divorce status, self-reported health status, financial satisfaction, level of happiness, and additional variables. The data-set contains 38,960 rows, providing ample opportunities to analyze the relationships between these variables and gain valuable insights into the factors that could potentially impact an individual's overall life satisfaction.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-var
#| tbl-cap: First five rows of the dataset, for illustrative purposes

data %>%
  select(1:6) %>%
  slice(1:5) %>%
  kable(
    col.names = c(
      "Year",
      "Sex",
      "No. of children",
      "Self-reported health",
      "Financial satisfaction",
      "Happiness"
    ),
    booktabs = TRUE,
    linesep = ""
  )
```

-   "year": This variable indicates the year in which the data was collected from GSS. In this data-set, all the observations are from the year of 1972 to 2012.
-   "sex": Categorical variable with two levels: male and female.
-   "babies": Integer variable with different levels, depending on the number of babies, for instance : 0, 1, 2, 3, 4, or 5+. For analysis purposes we construct "babies_number". The babies_number variable is an ordinal variable denoting the number of babies an individual in the dataset has. It ranges from 0 to 4, where 0 represents "0" babies, 1 denotes "1" baby, 2 signifies "2" babies, 3 indicates "3" babies, and 4 refers to "4" babies. The value 5 is reserved for any unclassified or missing responses.
-   "health": This variable indicates the self-reported health status of the individual at the time. It is a categorical variable with different levels, depending on the health status, for example : excellent, good, fair, or poor. We also use an analogous "health_score", which is an ordinal variable shows the self-reported health status of the individuals, which sign the ranges from 1 to 4. With 1 denoting "poor" health, 2 signifying "fair" health, 3 representing "good" health, and 4 referring to "excellent" health. The value 5 is reserved for any unclassified or missing responses.
-   "finance_satisfied": This variable indicates the individual's different level of satisfaction with their current financial situation. It is a categorical variable with different satisfaction levels: very satisfied, more or less satisfied, not very satisfied, and not at all satisfied. We also develop the analogous "finance_score". This ordinal variable represents the financial satisfaction level of the individuals in the dataset. It ranges from 1 to 3, with 1 indicating "Not at all satisfied," 2 denoting "more or less satisfied," and 3 signifying "pretty well satisfied." The value 4 is reserved for any unclassified or missing responses.
-   "happy": This variable indicates the individual's level of happiness or life satisfaction. It is a categorical variable with different levels of happiness: very happy, pretty happy, not too happy. We create an analogous happiness score ""happy_score". The happy_score variable is an integer that represents the individuals in the dataset's self-reported happiness levels. Its sign ranges from 0 to 3, with 0 denoting "not too happy," 1 denoting "happy," 2 denoting "very happy," and 3 holding any unclassified or missing responses.

@@

@fig-one explores the relationship between the number of children in a family and happiness levels, categorized as 'very happy,' 'pretty happy,' and 'not too happy.' Although childless families show the highest numbers in each happiness category, the data collected for such families is more abundant, necessitating an analysis of happiness proportions within each family size. The findings reveal no linear correlation between the number of children and overall family happiness, with happiness varying across family sizes without a consistent pattern. The non-uniform distribution of happiness categories among families with different numbers of children implies that factors beyond the number of children also contribute to happiness levels. In conclusion, the data shows variations in happiness levels among families of different sizes but does not establish a definitive correlation between the number of children and overall family happiness. Further research is needed to better understand the complex factors influencing happiness within families.

```{r}
#| label: fig-one
#| fig-cap: Relationship between Number of Babies and Happiness
#| echo: false
#| warning: false

data %>%
  group_by(babies, happy) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(babies), y = prop, fill = happy)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(name = "Number of Babies", expand = c(0.1, 0.1)) +
  scale_y_continuous(labels = scales::percent_format(), name = "Percentage", expand = c(0.05, 0)) +
  scale_fill_discrete(name = "Happiness") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
        plot.subtitle = element_text(size = 14, margin = margin(b = 20)),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")
```

@fig-two illustrates how gender distribution varies depending on the number of babies in the family. By comparison, a higher percentage of females than males is consistently observed across all groups. On the other hand, the lowest difference in sex distribution is found among families with no babies, where the female-to-male ratio is relatively closer. As the number of babies in a family increases, the percentage of females increases, with the highest difference observed in families with one to four babies. In these groups, the sex distribution is more skewed towards females.

```{r}
#| label: fig-two
#| fig-cap: Relationship between Number of Babies and Gender
#| echo: false
#| warning: false
data %>%
  group_by(babies, sex) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(babies), y = prop, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(name = "Number of Babies", expand = c(0.1, 0.1)) +
  scale_y_continuous(labels = scales::percent_format(), name = "Percentage", expand = c(0.05, 0)) +
  scale_fill_discrete(name = "Sex") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
        plot.subtitle = element_text(size = 14, margin = margin(b = 20)),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        legend.position = "bottom")
```

@fig-three shows analysis and visualization of the number of children born each year between 1972 and 2012 in connection to the degree of happiness to offer a more in-depth and complete perspective of the data. In 1974, 1985, 1993, 2004, 2006, and 2008, a relatively high percentage of people reported feeling unpleasant concerning the variable "not too happy." On the other hand, the proportion of people reporting being cheerful in 1973, 1977, 1994, 2006, and 2012 was relatively high. The data shows that excellent, fair, and sound are the most prevalent self-rated health statuses among adults with one or two children. It cannot be claimed that the number of children will impact the health scores, even though participation rates in families with one kid are generally high, even as time goes on.

@@

ggplot(data[data$happy == "very happy", ], aes(x = year, fill = factor(babies))) +
  geom_bar(position = "fill") +
  facet_wrap(~health, nrow = 5, scales = "free_y") +
  labs(x = "Year", y = "Percentage of Babies", fill = "Number of Babies") +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0.05, 0)) +
  theme(panel.grid.major.x = element_line(color = "grey", size = 0.1),
        strip.text = element_text(size = 10),
        panel.spacing = unit(0.2, "lines"))
```

@@

@fig-four shows that when individuals have one child, more participants rate their health as 'good' and 'excellent' compared to other self-health rating categories. By contrast, for the individuals who do not have children, the proportion of those who rate their health as "poor" is the highest; however, no clear conclusion can be drawn from this finding. For individuals with three children, the feedback on their health status is evenly distributed among "good," "fair," and "excellent," with a tiny proportion of participants reporting "poor" health. Overall, having one child is associated with most participants reporting an excellent health status.

@@

@fig-five illustrates the association between health and the number of babies in the family is investigated via a dot plot. The information is divided into five categories depending on the number of babies, which from 0 to 4, and four categories based on health, including "excellent," "fair," "good," and "poor." from the year of 1970 to the 2010. The graph shows a trend that may point to a link between a family's health condition and the number of children they have. By comparison; with more kids are born to families with "excellent" health than to those with 'poor' health, with resulting in a more balanced distribution of births. As the result; this pattern suggests that having one or more children is more likely when one is in better health self-rate. The percentages for all four health categories tend to decline as the number of newborns rises. This trend is most apparent in the "poor" health category, where the percentage drop is more dramatic as the number of births rises. The information points to a connection between family size and health. The chance of having one or more children is often more significant when one is in better health than when one is in worse health. But it's important to remember that the link could not be direct and that other variables might also affect how many children a family has.

@@

ggplot(data[data$happy == "very happy", ], aes(x = year, y = factor(babies))) +
  geom_point(position = position_jitter(width = 0.2, height = 0.1), alpha = 0.3, size = 2, color = "#4daf4a") +
  facet_wrap(~health, ncol = 5, scales = "free_y") +
  labs(x = "Year", y = "Number of Babies") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.grid.major.x = element_line(color = "grey", size = 0.1),
        strip.text = element_text(size = 10),
        panel.spacing = unit(0.2, "lines"))
```

@@

@fig-six suggests the connection between the number of babies and self financial satisfaction rate. Regardless of the group, the less financially satisfied families having one or more children.On the other hand; most families in all financial satisfaction levels have 0 children, which is intriguing. In contrast to the more or less content and 'not satisfied at all' categories, the proportion of families with no children is much higher in the 'very well satisfied' group. As the number of babies grows, the percentages for all three types of financial satisfaction tend to decrease. This trend is evident in the "pretty well satisfied" category, where the percentage decrease is steeper as the number of babies increases. In conclusion, the results point to a relationship between family size and financial contentment. Families that are more financially satisfied tend to have fewer children, whereas those who are less financially satisfied tend to have more. Remember that this link could not be linear and that other variables might affect how many children a family has.

```{r}
#| label: fig-six
#| fig-cap: Relationship between Happiness and finance satified
#| echo: false
#| warning: false

ggplot(data, aes(x = finance_satified, fill = factor(babies))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Rate of Finance Satisfaction",
       y = "Percentage of Cases by Number of Babies",
       fill = "Number of Babies") +
  theme_minimal() +
  coord_flip()
```

# Model

We use linear regression to estimate the following equation:

@@

In conclusion, the data analysis presented in @fig-one to @fig-six reveals several key findings related to the number of children in a family and its association with happiness, health, and financial satisfaction. Overall, the families without children reported that the highest happiness levels, however; the larger sample size of childless families could influence this result. The data on gender distribution shows that there are more women in families with more children compared with male. The most significant differences between families with one child and those with four children can be seen.

@@

## Relationship between the number of children and their gender.

@tbl-sex indicates that the sex distribution varies depending on the number of babies in the family. Across all groups, there is a consistent pattern of a higher percentage of females compared to males. The lowest difference in sex distribution is observed among families with no babies, where the female-to-male ratio is relatively closer 54.53% female vs. 45.47% male. As the number of babies in a family increases, the percentage of females also increases, with the highest difference observed in families with 1 and 4 babies. In these groups, the sex distribution is more skewed towards females, with 60.92% and 60.98% female representation, respectively.

```{r}
#| label: tbl-sex
#| tbl-cap: Relationship between Number of Children and Gender
#| echo: false
#| warning: false

sex_table <- data %>% 
  group_by(babies, sex) %>% 
  summarize(response = n(), variance = var(as.numeric(sex))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, sex, percent) %>%
  spread(sex, percent)

kable(sex_table  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Happiness between families with different numbers of babies

@tbl-happ suggests that families with 4 babies have the highest proportion of "happy" individuals at 63%. In contrast, families with 3 babies have the highest proportion of "not too happy" individuals at 16% and the lowest proportion of "happy" individuals at 50%. Families with 0, 1, and 2 babies have relatively similar distributions of happiness levels. It is worth noting that families with 3 babies have the highest proportion of "very happy" individuals at 33%, followed closely by families with 2 babies at 32%. Families with 4 babies have the lowest proportion of "very happy" individuals at 24%.

```{r}
#| label: tbl-happ
#| tbl-cap: Happiness Levels Across Families with number of Babies
#| message: false
#| echo: false
#| warning: false

happy_table <- data %>% 
  group_by(babies, happy) %>% 
  summarize(response = n(), variance = var(as.numeric(happy))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, happy, percent) %>%
  spread(happy, percent)

kable(happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Self-rated health and the number of babies

In @tbl-heal, which appears suggests that individuals without any babies tend to have higher percentages in all health categories, especially in the "poor" health category (93%). As the number of babies in a family increases, the proportion of individuals in the "excellent," "fair," and "good" health categories generally decreases. The "poor" health category also follows this trend, but the decrease is less pronounced. Individuals with 1 or 2 babies have relatively similar health status distributions, with slightly higher percentages in the "excellent" and "good" health categories compared to the "fair" and "poor" health categories. Individuals with 3 or 4 babies have the lowest percentages across all health categories.

```{r}
#| label: tbl-heal
#| tbl-cap: Health Status with number of Babies
#| message: false
#| echo: false
#| warning: false

health_happy_table <- data %>%
  group_by(health, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(health, babies, percent) %>%
  spread(babies, percent)

kable(health_happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

@@

@tbl-fin suggests that families with 0 babies have the highest percentages of financial satisfaction across all categories. As the number of babies in a family increases, the proportion of families in each financial satisfaction category generally decreases. Families with 4 babies have the lowest percentages in all financial satisfaction categories.Notably, the "pretty well satisfied" category shows the steepest decline in percentages as the number of babies increases.Families with 0 babies in this category have the highest percentage of financial satisfaction at 85.47%, while families with 4 babies have the lowest percentage at 0.05%.

@@

The analysis of the GSS data showed in @fig-one and @fig-five, which suggested the relationship between the number of children in a family and the level of the happiness and health status of individuals. In compare; the families with no children seem to have the highest proportion of 'very happy', 'pretty happy' responses rate. Although this could be influenced by the larger amount of data collected for childless individuals. On the other hand, the families with 'excellent' health tend to have a more balanced distribution of children, with a higher likelihood of having one or more children compared to families with poorer health.

In addition, the relationship between family size and happiness, as well as family size and health. The data analysis in @fig-two also indicated the relationship between family size and gender distribution. This paper suggested that the sex distribution varies depending on the number of babies in the family, with a higher percentage of females than males observed across all groups. As the number of babies in a family increases, at the same time, the percentage of females also increases, with the highest difference observed in families with one and four babies. By comparison, in these groups, the gender distribution is more skewed towards females. This finding suggested that the presence and number of babies in a family may also influence the gender composition of the family.

On the other hand, the data from @fig-three and @fig-four supports the relationship between family size, happiness, and health. The analysis of the number of children each year from 1972 to 2012 about the level of happiness shows the proportion of individuals who reported feeling "not too happy" and "happy" over the years. This finding highlights the potential influence of horticultural and economic factors on happiness and family size over time.

Additionally, the data from @fig-four shows that families with one child are associated with the highest proportion of participants reporting 'excellent' health status. This finding might suggest that having a smaller family size could be related to better health outcomes for individuals. However, it is crucial to consider other factors that could influence the health status of individuals, such as access to healthcare, lifestyle choices, and overall well-being.

@@

## Weakness

@@

## Next Steps

@@

NAME : SHAOHAN CHANG Institution/Department : Department of Statistical Sciences (University of Toronto) Email Address : shaohan.chang\@utoronto.ca
